ropenblas: Download, Compile and Link OpenBLAS Library with R

### 2020-10-01 paper/paper.Rmd
Reducing the article from 13 to 9 pages

```{r logo, echo = FALSE, message = FALSE, fig.cap = "Computer library logo.", fig.align = "center", out.width="30%"}
knitr::include_graphics(path = "logo_ropenblas.png")
```

@@

These programs that are described in `SystemRequirements` in the package's `DESCRIPTION` file are essential for compiling the OpenBLAS library and the R language. The functions of the `ropenblas` package are designed to identify the lack of these dependencies external to CRAN, informing the package user which dependencies are missing and suggesting that they should be installed. The other dependencies indexed to CRAN are described in `Imports` in the file  `DESCRIPTION`. These will be installed automatically.

Other warnings can also be suggested, such as, for example, a problem with the internet connection. All warnings are given very clearly so that the user has no doubts about the problem that may be occurring.

@@

```{r, prompt=TRUE, dependson="ropenblas", cache=TRUE}
last_version_r(major = 3L)
```

Note that the `last_version_r()` function returns a list of three elements, where the first element called `last_version` returns the latest version of the R language based on the argument `major`. Thus, the latest version of the R language since the largest version is $3$ is version $3.6.3$. The second element of name `versions` of the return list is a vector containing all versions of the language R for `major = 3L`. The third and last element of the return list is named `n` and returns an integer value referring to the number of versions of R given `major = 3L`. By default, the `major` argument of the `last_version_r()` function is `NULL`. Considering the standard, the function will always consider the largest current version of the language, as in the example below:

```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE}
last_version_r(major = NULL)
```

@@

```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE}
last_version_openblas()
```

@@

After executing the function, everything will be ready, and can be checked by executing the command `sessionInfo()` in the next execution of R, that is, in a new instance. Briefly, the `sessionInfo()` command will return something like:

```{r, prompt=TRUE}
sessionInfo()
```

Thus the user will know that the compilation of the R library, and if necessary, of the OpenBLAS library was carried out as requested. The user can also use the command `extSoftVersion()["BLAS"]` if it is only necessary to return which version of the OpenBLAS library is being considered, according to the following code:

@@

@@

**Example**: Assuming that the user of the package has already used the function `rcompiler(x = "4.0.0")` and is in a version other than 4.0.0 but intends to return to version 4.0.0, he can return without compile the language again, as shown in the image below. The output was placed in the form of an image since the return of the functions of the package use symbols that facilitate the user of the functions of the package to understand the suggestions, as well as understand if something went right or wrong.

```{r, eval=FALSE}
rcompiler(x = "4.0.0")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="A possible exit from the 'rcompiler' function. Suggestion of switching between versions of R without the need to recompile the language.", out.width="100%"}
knitr::include_graphics("linkar_again.png")
```

Due to the internal dependencies of the `ropenblas` package, it is advisable not to try to go to versions of R prior to 3.1.0, as these dependencies do not work on these versions of R. Therefore, in earlier versions of R, installing the `ropenblas` package will not be possible. Therefore, if the user wishes to return to a more current version of R, he must install the language R in the GNU/Linux distribution without the help of the package. Although it is allowed to install previous versions of R prior to 3.1.0, it will only be done after the user responds to some questions of an alert message, as shown in the image below:

```{r, eval=FALSE, prompt=FALSE}
rcompiler(x = "3.0.0")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Attempted to install an R language version prior to version 3.1.0.", out.width="100%"}
knitr::include_graphics("attention.png")
```

@@

The code below is a small example of the benefits of considering linking the R language to the OpenBLAS library, in which singular-value decomposition of a rectangular matrix is computed, codes that are executed on the same machine, version of R and section. The first part of the code was executed with R without being linked to a version of the OpenBLAS library and the second part was executed after using the `ropenblas()` function. Note that the code when executed in R linked to a version of the OpenBLAS library can be clearly more efficient:

```{r, eval = FALSE}
# Using the BLAS + Lapack library:
# Matrix products: default
# BLAS:   /usr/lib/libblas.so.3.9.0
# LAPACK: /usr/lib/liblapack.so.3.9.0
n <- 1e3L
X <- matrix(rnorm(n * n), n, n)
system.time(svd(X)) 
## user      system      elapsed
## 7.246     0.029       7.286

# ----

# After linking the OpenBLAS library to R using the ropenblas() function:
# Using the OpenBLAS version 0.3.10 library:
# Matrix products: default
# BLAS/LAPACK: /opt/OpenBLAS/lib/libopenblas_haswellp-r0.3.10.so
n <- 1e3L
X <- matrix(rnorm(n * n), n, n)
system.time(svd(X)) 
## user      system    elapsed
## 3.024     2.004     0.677 
```

@@

Running the `link_again()` function in a situation where there is no need will not generate problems. The function will return the message that everything is linked correctly, according to the code and image that follows:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If relinking is not required, the function will make it clear.", out.width="100%"}
knitr::include_graphics("link_again_02.png")
```

>>>>

<!-- ```{r logo, echo = FALSE, message = FALSE, fig.cap = "Computer library logo.", fig.align = "center", out.width="30%"} -->
<!-- knitr::include_graphics(path = "logo_ropenblas.png") -->
<!-- ``` -->

@@

<!-- These programs that are described in `SystemRequirements` in the package's `DESCRIPTION` file are essential for compiling the OpenBLAS library and the R language. The functions of the `ropenblas` package are designed to identify the lack of these dependencies external to CRAN, informing the package user which dependencies are missing and suggesting that they should be installed. The other dependencies indexed to CRAN are described in `Imports` in the file  `DESCRIPTION`. These will be installed automatically. -->

<!-- Other warnings can also be suggested, such as, for example, a problem with the internet connection. All warnings are given very clearly so that the user has no doubts about the problem that may be occurring. -->

@@

```{r, prompt=TRUE, dependson="ropenblas", cache=TRUE}
last_version_r(major = 4L)
```

<!-- Note that the `last_version_r()` function returns a list of three elements, where the first element called `last_version` returns the latest version of the R language based on the argument `major`. Thus, the latest version of the R language since the largest version is $3$ is version $3.6.3$. The second element of name `versions` of the return list is a vector containing all versions of the language R for `major = 3L`. The third and last element of the return list is named `n` and returns an integer value referring to the number of versions of R given `major = 3L`. By default, the `major` argument of the `last_version_r()` function is `NULL`.  -->

<!-- Considering the standard, the function will always consider the largest current version of the language, as in the example below: -->

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->
<!-- last_version_r(major = NULL) -->
<!-- ``` -->

@@

<!-- ```{r, dependson="ropenblas", cache=TRUE, prompt=TRUE} -->
<!-- last_version_openblas() -->
<!-- ``` -->

@@

After executing the function, everything will be ready, and can be checked by executing the command `extSoftVersion()["BLAS"]` in the next execution of R, that is, in a new instance. Thus the user will know that the compilation of the R library, and if necessary, of the OpenBLAS library was carried out as requested. Briefly, the `extSoftVersion()["BLAS"]` command will return something like:

@@

<!-- ```{r, prompt=TRUE} -->
<!-- sessionInfo() -->
<!-- ``` -->

@@

<!-- **Example**: Assuming that the user of the package has already used the function `rcompiler(x = "4.0.0")` and is in a version other than 4.0.0 but intends to return to version 4.0.0, he can return without compile the language again, as shown in the image below. The output was placed in the form of an image since the return of the functions of the package use symbols that facilitate the user of the functions of the package to understand the suggestions, as well as understand if something went right or wrong. -->

<!-- ```{r, eval=FALSE} -->
<!-- rcompiler(x = "4.0.0") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="A possible exit from the 'rcompiler' function. Suggestion of switching between versions of R without the need to recompile the language.", out.width="100%"} -->
<!-- knitr::include_graphics("linkar_again.png") -->
<!-- ``` -->

<!-- Due to the internal dependencies of the `ropenblas` package, it is advisable not to try to go to versions of R prior to 3.1.0, as these dependencies do not work on these versions of R. Therefore, in earlier versions of R, installing the `ropenblas` package will not be possible. Therefore, if the user wishes to return to a more current version of R, he must install the language R in the GNU/Linux distribution without the help of the package. Although it is allowed to install previous versions of R prior to 3.1.0, it will only be done after the user responds to some questions of an alert message, as shown in the image below: -->

<!-- ```{r, eval=FALSE, prompt=FALSE} -->
<!-- rcompiler(x = "3.0.0") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="Attempted to install an R language version prior to version 3.1.0.", out.width="100%"} -->
<!-- knitr::include_graphics("attention.png") -->
<!-- ``` -->

@@

<!-- The code below is a small example of the benefits of considering linking the R language to the OpenBLAS library, in which singular-value decomposition of a rectangular matrix is computed, codes that are executed on the same machine, version of R and section. The first part of the code was executed with R without being linked to a version of the OpenBLAS library and the second part was executed after using the `ropenblas()` function. Note that the code when executed in R linked to a version of the OpenBLAS library can be clearly more efficient: -->

<!-- ```{r, eval = FALSE} -->
<!-- # Using the BLAS + Lapack library: -->
<!-- # Matrix products: default -->
<!-- # BLAS:   /usr/lib/libblas.so.3.9.0 -->
<!-- # LAPACK: /usr/lib/liblapack.so.3.9.0 -->
<!-- n <- 1e3L -->
<!-- X <- matrix(rnorm(n * n), n, n) -->
<!-- system.time(svd(X))  -->
<!-- ## user      system      elapsed -->
<!-- ## 7.246     0.029       7.286  -->

<!-- # ---- -->

<!-- # After linking the OpenBLAS library to R using the ropenblas() function: -->
<!-- # Using the OpenBLAS version 0.3.10 library: -->
<!-- # Matrix products: default -->
<!-- # BLAS/LAPACK: /opt/OpenBLAS/lib/libopenblas_haswellp-r0.3.10.so -->
<!-- n <- 1e3L -->
<!-- X <- matrix(rnorm(n * n), n, n) -->
<!-- system.time(svd(X))  -->
<!-- ## user      system    elapsed -->
<!-- ## 3.024     2.004     0.677  -->
<!-- ``` -->

@@

<!-- Running the `link_again()` function in a situation where there is no need will not generate problems. The function will return the message that everything is linked correctly, according to the code and image that follows: -->

<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.cap="If relinking is not required, the function will make it clear.", out.width="100%"} -->
<!-- knitr::include_graphics("link_again_02.png") -->
<!-- ``` -->
